### 2019-04-07 fig1_2app.Rmd

---
title: "supplement"
author: "Selcen Arı"
date: "7 Nisan 2019"
output: html_document
---

```{r}
data("minsamp")

minsamp
```

```{r}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
  update_nodes(once=TRUE)%>%
  update_how("Gene2", 2)%>%
  update_nodes()%>%
  simulate(3)%>%activate(edges)%>%
  as_tibble()%>%
  select(Competing_name,comp_count_list,effect_list)%>%
  unnest()

```

```{r}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
  update_nodes(once=TRUE)%>%
  update_how("Gene2", 2)%>%
  update_nodes()%>%
  simulate(3)%>%activate(edges)%>%
  as_tibble()%>%
  select(Competing_name,comp_count_list,effect_list)%>%
  unnest()
```