### 2019-08-09 article.tex
overview

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Network based multifactorial modelling microRNA-target interaction.},
            pdfauthor={Selcen Ari; Alper Yilmaz},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Network based multifactorial modelling microRNA-target interaction.}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Selcen Ari\footnote{Department of Bioengineering, Yildiz Technical
  University, Istanbul, Turkey} \\ Alper Yilmaz\footnote{Department of Bioengineering, Yildiz Technical
  University, Istanbul, Turkey}}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{09 Ağustos 2019}

\begin{document}
\maketitle

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

Competing endogenous RNA (ceRNA) regulations and crosstalk between
various types of non-coding RNA in human is remarkable in means of miRNA
regulation. Many studies have pointed out that an alteration in
miRNA:target interaction can result in unexpected changes due to
indirect and complex interactions. In this paper, we defined a new
network-based model that handles miRNA:ceRNA interactions with
expression values. Our model is able to handle miRNA interaction factors
such as seed type, binding energy, if provided. Our approach is able to
reveal that a perturbation in an element of network affects whole
competing elements differently and cooperative efficiencies of miRNAs on
common targets could be calculated. Our findings emphasized importance
of miRNA:target ratios being crucial, as reported by previous studies.
We have showed that the competing elements which have the same or close
expression values may not be affected equally from the perturbation
because of repression functionality depended on interaction factors of
miRNA target pairs (TOFIX: Hocam etkileşim faktörlerinden dolayı
ekspresyonları yakın olsa bile baskılanma aktitesinin aynı
sergilenmeyeceğini ifade etmeye çalıştım. Bu şekilde anlaşılıyor mu?).
We applied the model to real sample consisting of breast cancer gene and
miRNA expression dataset and experimental miRNA:target interaction
dataset all generated via high throughput sequencing methods. A gene
over-expressed in tumor tissue, namely \emph{ABCC1}, is used as
perturbing element. We have observed that change in expression level of
single gene in miRNA:target network is sufficient to perturb regulations
in whole network, due to unforeseen and unpredicted regulation which are
only visible when considered in network context. Therefore, this model
helps unveiling the crosstalk between elements in miRNA:target network
where abundance of target and sponge effect are taken into account. The
model is scalable and can be plugged in with emerging miRNA effectors
such as circRNAs. The model is available as R package
\href{https://github.com/selcenari/ceRNAnetsim}{ceRNAnetsim}.

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

MicroRNAs (miRNAs) are a family of short non-coding RNAs which are key
regulator of gene expression through various post-transcriptional
mechanisms. Although the mechanisms by which miRNA represses are not
fully understood, miRNAs predominantly repress their targets. Repressive
activities of miRNAs vary depending on many factors that are significant
to microRNA:target interactions. These factors include microRNA:target
binding energy, binding location in target sequence, base pairing types
between miRNA and target, abundance of miRNAs and targets (Grimson et
al. 2007). Binding energies of miRNA:target complexes vary based on
nucleotide context and determine folding stability of complex (Cao and
Chen 2012). It has been demonstrated that the binding energy between
miRNA and target indicates stability or affinity of complex (Helwak et
al. 2013) and does not directly determine repressive activity of miRNA
(Cao and Chen 2012). Early studies have argued that 2-8 nt sequence,
seed, located in miRNA 5'end bind to specific sequence located in 3'UTR
of its target (Bartel 2004; Lewis, Burge, and Bartel 2005). In recent
studies, it has been shown that miRNAs can interact with targets via
sequences located in regions such as 5'UTR or CDS (Hausser et al. 2013;
Helwak et al. 2013; Moore et al. 2015). These studies also showed that
binding location could indicate functionality of miRNA:target
interaction or be effective on abundance of targets(TOFIX abundance
etkisi derken, bazen azaltabilir, bazen arttırabilir, manasında mı?,
evet hocam anlaşılmıyorsa değiştirebilirim). It has been shown that
miRNAs exhibit repressive activity via, 6-8 nt long sequence that is
perfectly complementary with targets, seed region at the 5' end of
miRNAs (Bartel 2009; Grimson et al. 2007). On the other hand, some
researchers have reported that seed sequence of miRNA can have
mismatches or bulged/wobble nucleotides and may locate in region other
than 5'end of miRNAs (Hafner et al. 2010; Helwak et al. 2013). On top of
all these factors, abundance of miRNAs and targets and miRNA:target
ratio in cells predominantly affect efficiency of miRNA:target
interaction (Arvey et al. 2010; Bosson, Zamudio, and Sharp 2014; Denzler
et al. 2014).

As it is possible for miRNAs to suppress multiple targets, an individual
mRNA molecule can also be targeted by multiple miRNAs. In that case, the
targeted mRNAs exhibit competitor behavior, that is hypothesized as
competing endogenous RNAs (ceRNAs) (Ala et al. 2013; Cesana and Daley
2013), against their miRNAs. Briefly, Ala et al.~have explained the
ceRNA hypothesis as disturbance of the other target when one of the
targets on a steady-state system that included one miRNA and two target
was perturbed with expression change (Ala et al. 2013). Regarding
interaction between miRNAs and their target in a cell, explaining and
predicting results of an individual perturbation is difficult due to
complexity of interactions. Various computational and experimental
studies have tackled the problem of unraveling ceRNA:miRNA interactions.
It has been observed that when abundance of one of the targets of
miR-122 was increased, the other target expression also slightly
increased as a result of decreasing repressive activity of miR-122 on
them (Denzler et al. 2014). Bosson et al.~have developed a mathematical
model for changes on total target pool concentration after grouping
targets according to affinity and demonstrated that miRNA activity
correlated with affinity between miRNA and target (Bosson, Zamudio, and
Sharp 2014). Cooperative efficiency of miRNAs as well as competitor
behaviors of targets were also studied and it has been demonstrated to
be crucial for regulating available mRNA levels of targets (Denzler et
al. 2016). MiRNA:target interactions have been modeled as stoichiometric
and catalytic mechanisms and Figliuzzi et al.~have recommended handling
models in network context (Figliuzzi, Marinari, and De Martino 2013).
The model that can explain miRNA target interaction through topological
features has been applied at bipartite network by Nitzan et al.~(Nitzan
et al. 2014). Robinson and Henderson applied the model that handles
miRNA:target direct and indirect interactions via common miRNA of gene
and target of miRNA, at bipartite network. It has been demonstrated that
all miRNAs and targets in the network can interact with each other
through common miRNAs and genes, without interaction between the same
type of nodes (Robinson and Henderson 2018).Associated genes that are
targets of the same miRNAs have been found with help of correlation of
gene expression changes in recent algorithm (Markus List 2017). List et
al.~have specified that their approach can be useful for ceRNA studies
and published their approach as an R package.

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

\hypertarget{construction-of-mirnatarget-network}{%
\subsubsection{Construction of miRNA:target
network}\label{construction-of-mirnatarget-network}}

miRNA and target pairs per line should be provided as edge list to
construct the network. At each line minimum required information is
expression levels of miRNA and the target. If available, additional data
about factors effecting binding or efficiency of miRNA can be provided
as separate columns. After construction of the network, amount of miRNA
per target is calculated and kept as edge data. Simply, a target will
sequester miRNA proportional to its ratio amount among other targets. If
additional criteria effecting the binding of miRNA to its target is
provided, distribution of miRNA will be calculated accordingly. Target
can be mRNA or any other ceRNA (circRNA, ncRNA, etc.) thus, throughout
the manuscript terms target, gene and ceRNA are used interchangeably.

\hypertarget{triggering-perturbation-and-subsequent-calculations}{%
\subsubsection{Triggering perturbation and subsequent
calculations}\label{triggering-perturbation-and-subsequent-calculations}}

Initially, the network is assumed in steady-state (\autoref{fig1}a)
condition and needs a trigger for initiating calculations. The trigger
can be a change in expression level of one or more genes
(\autoref{fig1}b). After a trigger, the network undergoes iterative
cycle of calculations at each of which distribution of miRNA in local
neighborhood is recalculated (\autoref{fig1}c). Based on new miRNA
distribution, expression level of each node (i.e.~ceRNA) is updated. Due
to nodes which are targeted by multiple miRNAs, the change in one
neighborhood spreads to other neighborhoods (\autoref{fig1}d),
consequently have potential to effect whole network due to ``ripple
effect''.

\begin{figure}
\hypertarget{fig1}{%
\centering
\includegraphics{Fig1.jpg}
\caption{Schematic presentation of mechanism of network based model. a)
In steady state, miRNAs (M: gray triangle) repress targets (G: circle
coconut) according to proportion of their targets' expression. b) Two
fold increase in transcript level of Gene2 (G2) acts as a trigger. c)
Distribution of miRNA1 (M1) changes. d) The change at expression of
common target effects changes of proportional distribution of miRNA2
(M2). Expression values are rounded to integers for
simplicity}\label{fig1}
}
\end{figure}

During calculations, following assumptions were adopted; 1)
Transcription and degradation rates of miRNAs are steady and equal. 2)
All available miRNAs are recycled as in miRNA:ceRNA binding, target is
degraded and miRNA is unaffected. 3) ceRNA targets also have stable
transcription and degradation rates and these rates are equal.

The repression efficiency of a miRNA on the individual target
(\(Eff_{gi}\)) is calculated according to equation \eqref{eq:1}; where
miRNA expression (\(C_m\)) in local neighborhood is distributed among
targets using individual gene expression levels (\(C_{gi}\)) (\#TODO
isn't \(C_g\) gene expressions in the group is equal to total of
\(C_{gi}\) in group?). For the genes targeted by multiple miRNAs,
cooperative activity of miRNAs, \(R\), is calculated by summing
repression activity of each miRNA (Equation \eqref{eq:2}).

\begin{equation} 
    Eff_{gi}= C_m \times C_{gi}/\sum_{1}^{i} C_g \tag{1}\label{eq:1}
\end{equation} \begin{equation}
   R_{gi}= Eff_{i1} + Eff_{i2} ... \tag{2}\label{eq:2}
\end{equation}

\hypertarget{multifactorial-calculations-in-mirnatarget-network}{%
\subsubsection{Multifactorial calculations in miRNA:target
network}\label{multifactorial-calculations-in-mirnatarget-network}}

Interactions between miRNAs and their targets can be affected from
various factors. So, our model integrates multiple factors when
calculating overall miRNA activity. We classified factors into two
categories. Factors effecting binding determine interaction between
miRNA and target and they alter amount of miRNA sequestered to target.
Factors effecting efficiency determine proportion of target mRNA
degraded by bound miRNA (\#TODO fix this sentence). In other words,
binding factors exert their influence before or during binding,
efficiency factors exert their influence after binding. In the
literature, binding free energy (Cao and Chen 2012 ; Helwak et al. 2013)
and seed type (Werfel et al. 2017) in miRNA:target interactions are
described as factors effecting binding affinity. Efficiency factors
determine how many of miRNA:target complexes will result in inhibition
and binding region on target drastically effect miRNA degradation
efficiency (Hausser et al. 2013; Helwak et al. 2013). Both binding and
efficiency factors are normalized to their maximum values and scaled to
{[}0,1{]} interval. The normalized values of factors take into account
to determine binding activity and miRNA efficiency on targets
(\autoref{fig2}). Binding affinities (activity, \(Eff\)) of miRNAs on
each individual gene are calculated as shown in equation \eqref{eq:4};
\(C_m\), miRNA expression in the group; \(C_g\), Gene expression;
\(C_{gi}\), individual gene expression; gi, individual gene; g, whole
genes in a group (\autoref{fig2}c). \begin{equation}
Eff_{gi}= C_m \times E^\prime_{gi} \times STE^\prime_{gi} \times C_{gi}/(\sum_{1}^{i} E^\prime_{gi} \times STE^\prime_{gi} \times C_{gi}) \tag{4}\label{eq:4}
\end{equation} \begin{equation}
Eff_{gi}= Eff_{gi}*RE^\prime_{gi} \tag{5}\label{eq:5}
\end{equation}

\begin{figure}
\hypertarget{fig2}{%
\centering
\includegraphics{fig2.jpg}
\caption{Calculations to determine of miRNA binding and repression
efficiency. \(G\), Gene; \(M\), miRNA; \(STE\), seed type effect;
\(RE\), Region Effect; \(E\), Energy; \(STE^\prime\), normalized values
of seed type efficiency coefficient; \(RE^\prime\), normalized values of
region efficiency coefficient; \(E^\prime\), normalized values of energy
coefficient.}\label{fig2}
}
\end{figure}

\iffalse

\%\% (\#TODO the equation is not considering binding factors, right?. We
should explain the calculation made in simulate function here, in simple
sentences. the function looks like this: \%\%
input\_graph\%\textgreater{}\% \%\%
tidygraph::activate(edges)\%\textgreater{}\% \%\%
group\_by(to)\%\textgreater{}\% \%\% mutate(mirna\_count\_per\_comp =
mirna\_count\_current\emph{comp\_count\_current}afff\_factor/sum(comp\_count\_current\emph{afff\_factor),
mirna\_count\_per\_comp = ifelse(is.na(mirna\_count\_per\_comp), 0,
mirna\_count\_per\_comp))\%\textgreater{}\% \%\%
ungroup()\%\textgreater{}\% \%\% mutate(effect\_pre = effect\_current,
effect\_current =
mirna\_count\_per\_comp}degg\_factor)\%\textgreater{}\% \%\%
group\_by(from)\%\textgreater{}\% \%\% mutate(comp\_count\_pre =
ifelse(comp\_count\_current\textless{} 0, 1, comp\_count\_current),
comp\_count\_current = comp\_count\_pre -
(sum(effect\_current)-sum(effect\_pre)), comp\_count\_current =
ifelse(comp\_count\_current\textless{} 0, 1,
comp\_count\_current))\%\textgreater{}\% \%\%
ungroup()\%\textgreater{}\% \%\% mutate(comp\_count\_list =
pmap(list(comp\_count\_list, comp\_count\_current), c), effect\_list =
pmap(list(effect\_list, effect\_current), c), mirna\_count\_list =
pmap(list(mirna\_count\_list, mirna\_count\_current), c)) \%\% ) \fi

After miRNA binds to its target, but might not repress to bound target.
The functionality of bound miRNA on target depends on efficiency factors
like region that is binding sequence of miRNA on its target. Exact
repression efficiency of miRNA is calculated according to equation
\eqref{eq:5} (\autoref{fig2}d); \(RE^\prime_{gi}\), normalized values of
region efficiency coefficient between miRNA and gene. The cooperative
repression activity of miRNAs to their common targets is figured out as
shown in \autoref{fig2}e.

\hypertarget{breast-cancer-patient-dataset}{%
\subsubsection{Breast cancer patient
dataset}\label{breast-cancer-patient-dataset}}

We have applied our model in a real dataset for which experimental
measurements of various factors were available. Expression levels of
miRNA and genes in tumor and normal tissue of single patient are
retrieved from TCGA(\#TODO cite). High-throughput experimental datasets
which are provided miRNA:gene target pairs with interaction factors
(Helwak et al. 2013; Moore et al. 2015). We have combined miRNA and gene
expression datasets via miRNA:target gene dataset retrieved from
\ldots{} (\#TODO name of database and its
citation)-\textgreater{}suplementary datasets of (Helwak et al. 2013;
Moore et al. 2015) . Detailed description of network construction and
its code is available in Supplementary data (\#TODO burada link nasıl
olacak, word dosyası veya PDF dosyası ismi mi vermemiz mi gerekiyor,
yoksa sadece Sup Data denilmesi yeterli mi? ilgili dosya:
TCGA\_E9-A1N5\_article.Rmd). ABCC1 gene, over-expressed in tumor tissue,
was selected as trigger for simulation of integrated dataset. After
simulation of network, we have compared simulation results and tumor
tissue expression levels.

\hypertarget{results-and-conclusions}{%
\subsection{Results and Conclusions}\label{results-and-conclusions}}

\hypertarget{networks-with-single-factor}{%
\subsubsection{Networks with single
factor}\label{networks-with-single-factor}}

We have developed a network-based approach to assess effects of
expression level changes in competitive ceRNA regulation. The basic mode
of miRNA repression activity has been based on miRNA and target
abundance in various researches(Arvey et al. 2010; Denzler et al. 2014).
Our approach can effortlessly calculate effects of expression changes
when abundance levels of miRNAs and targets is only available factor. In
sample network given in Figure (\autoref{fig1}), after an increase in
expression level of a gene (G2), expression values of other genes also
changed due to redistribution of miRNA among its targets. Previous
studies have shown that if a gene abundance increases in ceRNA system,
expression levels of genes targeted by shared miRNA are also affected
(Lai, Wolkenhauer, and Vera 2016; Salmena et al. 2011; Tay, Rinn, and
Pandolfi 2014). It was observed that primary neighborhoods of the
trigger gene change with the same ratio, but cooperative efficiency of
miRNAs on common target causes differently change in common target
(\#TODO last sentence needs some clarification, after ``but
cooperative\ldots{}'') (\autoref{fig1}d). Genes targeted by multiple
miRNAs act as a trigger for adjacent local neighborhood of targeting
miRNAs, causing changes in expression levels of genes outside the local
neighborhood of original trigger gene. Therefore, primary expression
level change in gene (G2) causes changes in other group of genes (G5 and
G6) even though original trigger gene (G2) and genes in other group are
not targeted by common miRNA. In addition, as shown by ceRNA hypothesis
model of Ala et al., after the increase of gene expression level of G2,
the miRNA that is found in the same group (M1) tended to be less
repressive on its remaining targets (G1, G3 and G4). It's important to
note that the changes in gene expression levels will have more
pronounced effect if miRNA:target ratio is high, i.e., more miRNA
available per target, which was reported in previous findings (Arvey et
al. 2010; Bosson, Zamudio, and Sharp 2014; Denzler et al. 2014).
\href{}{}

\begin{figure}
\hypertarget{fig3}{%
\centering
\includegraphics{Fig3.jpg}
\caption{Target regulations with interaction parameters. a) In the
steady-state the repression activity of miRNAs on the targets after
binding and repression efficiency. b) The changes the repression
activities after increasing of G2 expression. c) Perturbation of primary
neighborhoods of M1 miRNA (M1 miRNA group). d) Regulation of gene
expression of other gene group via triggering target (common target
between M1 and M2).}\label{fig3}
}
\end{figure}

\hypertarget{competing-endogenous-rnatarget-networks-based-on-interaction-factors}{%
\subsubsection{Competing Endogenous RNA:Target networks based on
interaction
factors}\label{competing-endogenous-rnatarget-networks-based-on-interaction-factors}}

\href{}{\#TODO}: \textless{}\textgreater{} ``title should be more
informative''-\textgreater{}

Earlier studies reported that miRNA regulatory interactions are affected
by different parameters. For example, Xu et al.~have investigated the
importance of seed pairing type between miRNAs and their targets and
target site location by using proteomics dataset (Xu, Wang, and Liu
2014). They have proposed that the features of binding between miRNA and
target can be critical for miRNA efficiency. In addition, binding energy
between miRNA and targets is a significant determinant for miRNA
efficiency and it has been reported that strength of miRNA:target
interactions is depended on binding energy of complexes (Breda et al.
2015). Similarly, another study revealed that affinity is correlated
with seed pairing of miRNA:target pairs and suggested that affinity is
correlated with number of canonical seed base pairing (Bosson, Zamudio,
and Sharp 2014). Therefore, we have integrated aforementioned
interaction parameters that could be useful for predicting miRNA
repression activity more accurately. The sample dataset used in Figure
\autoref{fig1} is recalculated with additional factors in effect
\href{}{} with same trigger, two fold increase in Gene2
(\autoref{fig3}). When the factors were taken into account in the
system, miRNA efficiencies varied as shown in \autoref{fig3}a. Although
the miRNA:target expression ratios in steady-state were same in
comparison with the sample dataset without factors, efficiency of
binding and repression have changed (compare miRNA distribution values
between Figure \autoref{fig1}a and Figure \autoref{fig3}a). When
expression of Gene2 (G2) increased, expression values of all genes also
changed differentially because of contribution of efficiency factors
(\autoref{fig3} b,c,d). We have considered that in our approach energy
and seed type of pairs is significant for binding and targeted region is
important for repression. \href{}{}

knitr::kable(minsamp)

In regulation of miRNA target sample system (\autoref{fig3}), miRNA (M1)
repressive efficiency on the primary triggering gene (G2) is low in
steady-state. So it has been observed that the regulatory activities of
miRNA efficiencies on the targets are weak after the increase of miRNA
target G2. When model is triggered with two fold increase in expression
level of common target (G4), the changes of other gene expressions have
observed more prominent \href{}{}. Furthermore, change in expression
level of target gene that has strong miRNA repression efficiency
resulted in evident perturbation in network. On the other hand, it was
observed that Gene2 was weakly affected from change of Gene4 expression
because of its weak interaction factors. Since perturbation efficiency
of each gene is different, we have developed a function which screens
each gene in the network for their perturbation efficiencies.
\textbf{(see details in Supplementary File(fig1\_2app.Rmd)}. When we
applied the method on the minimal dataset, Gene4 has been found to be
the most efficient element in terms of number of perturbed elements and
miRNA2 (M2) has been found to be causing the highest mean in expression
changes. \href{}{}

\href{}{} We used our model to simulate a real dataset which contains
thousands of genes and hundreds of miRNAs. Our model can successfully
simulate perturbations in such large network despite complex behaviors
and struggle to reach steady-state \textbf{(see details in Supplementary
File(TCGA\_E9-A1N5\_article.Rmd)}. Simulations show that change in
expression level of single gene has potential to effect whole network,
perturbing almost all nodes. These observations are in accordance with
competing endogenous RNA hypothesis where genes targeted with many
common miRNAs subsequently transmit perturbation to neighboring groups.

Network based approaches for analyzing miRNA:target interactions have
been developed in earlier studies. An initial attempt demonstrated the
ceRNA crosstalk in a network-like minimal interaction structure with
concentrations of ceRNA and miRNAs (Figliuzzi, Marinari, and De Martino
2013). Next, a network based kinetic model integrating miRNA and target
rates of transcription, degradation, binding and unbinding was developed
(Nitzan et al. 2014) using high throughput experimental dataset about
miR-92a depletion (Helwak et al. 2013). In this study, it was
demonstrated that distant ceRNAs can interact with each other via
indirect links, and the interactions are effected depending on distance
between ceRNAs or topological features of network (Nitzan et al. 2014).
More recently, an approach to detect ceRNA interaction by using the
miRNA expression, gene expression and common miRNAs between gene targets
was developed (Markus List 2017) which was effective in analyzing genes
through miRNAs. Based on the observations that a miRNA can exhibit
strong functionality to a target but may not against an other, the
authors have concluded that common miRNA based approach may not be
suitable for understanding regulations of ceRNA interactions.

In our approach, we have not taken into account transcription,
degradation or binding rates of elements in network. Because, although
it is known as the miRNAs are highly stable, the transcription and
degradation rates of miRNAs change depending on cellular conditions
(Rüegger and Großhans 2012). However, due to lack of datasets including
degradation and transcription rates of miRNAs at specific cellular
conditions, such factors are neglected in our model. Additionally, other
regulation parameters such as gene-gene interactions and transcription
factors are ignored but the network structure is flexible and can
integrate additional regulation elements as they become available. In
the future, with developments in experimental techniques about
miRNAs:target interaction dynamica, more accurate results can be
obtained from our model resulting in better understanding and
predictions of abnormal regulations and pathways underlying diseases or
conditions.

\hypertarget{references}{%
\subsection*{References}\label{references}}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-ala_integrated_2013}{}%
Ala, U., F. A. Karreth, C. Bosia, A. Pagnani, R. Taulli, V. Leopold, Y.
Tay, P. Provero, R. Zecchina, and P. P. Pandolfi. 2013. ``Integrated
Transcriptional and Competitive Endogenous RNA Networks Are
Cross-Regulated in Permissive Molecular Environments.''
\emph{Proceedings of the National Academy of Sciences} 110 (18):
7154--9. \url{https://doi.org/10.1073/pnas.1222509110}.

\leavevmode\hypertarget{ref-arvey_target_2010}{}%
Arvey, Aaron, Erik Larsson, Chris Sander, Christina S Leslie, and Debora
S Marks. 2010. ``Target mRNA Abundance Dilutes microRNA and siRNA
Activity.'' \emph{Molecular Systems Biology} 6 (April).
\url{https://doi.org/10.1038/msb.2010.24}.

\leavevmode\hypertarget{ref-bartel_micrornas_2004}{}%
Bartel, David P. 2004. ``MicroRNAs.'' \emph{Cell} 116 (2): 281--97.
\url{https://doi.org/10.1016/S0092-8674(04)00045-5}.

\leavevmode\hypertarget{ref-bartel_micrornas:_2009}{}%
Bartel, David P. 2009. ``MicroRNAs: Target Recognition and Regulatory
Functions.'' \emph{Cell} 136 (2): 215--33.
\url{https://doi.org/10.1016/j.cell.2009.01.002}.

\leavevmode\hypertarget{ref-bosson_endogenous_2014}{}%
Bosson, Andrew D., Jesse R. Zamudio, and Phillip A. Sharp. 2014.
``Endogenous miRNA and Target Concentrations Determine Susceptibility to
Potential ceRNA Competition.'' \emph{Molecular Cell} 56 (3): 347--59.
\url{https://doi.org/10.1016/j.molcel.2014.09.018}.

\leavevmode\hypertarget{ref-breda_quantifying_2015}{}%
Breda, Jeremie, Andrzej J. Rzepiela, Rafal Gumienny, Erik van Nimwegen,
and Mihaela Zavolan. 2015. ``Quantifying the Strength of miRNA--Target
Interactions.'' \emph{Methods} 85 (September): 90--99.
\url{https://doi.org/10.1016/j.ymeth.2015.04.012}.

\leavevmode\hypertarget{ref-cao_predicting_2012}{}%
Cao, Song, and Shi-Jie Chen. 2012. ``Predicting Kissing Interactions in
microRNA--Target Complex and Assessment of microRNA Activity.''
\emph{Nucleic Acids Research} 40 (10): 4681--90.
\url{https://doi.org/10.1093/nar/gks052}.

\leavevmode\hypertarget{ref-cesana_deciphering_2013}{}%
Cesana, Marcella, and George Q. Daley. 2013. ``Deciphering the Rules of
ceRNA Networks.'' \emph{Proceedings of the National Academy of Sciences
of the United States of America} 110 (18): 7112--3.
\url{https://doi.org/10.1073/pnas.1305322110}.

\leavevmode\hypertarget{ref-denzler_assessing_2014}{}%
Denzler, Rémy, Vikram Agarwal, Joanna Stefano, David P. Bartel, and
Markus Stoffel. 2014. ``Assessing the ceRNA Hypothesis with Quantitative
Measurements of miRNA and Target Abundance.'' \emph{Molecular Cell} 54
(5): 766--76. \url{https://doi.org/10.1016/j.molcel.2014.03.045}.

\leavevmode\hypertarget{ref-denzler_impact_2016}{}%
Denzler, Rémy, Sean E. McGeary, Alexandra C. Title, Vikram Agarwal,
David P. Bartel, and Markus Stoffel. 2016. ``Impact of MicroRNA Levels,
Target-Site Complementarity, and Cooperativity on Competing Endogenous
RNA-Regulated Gene Expression.'' \emph{Molecular Cell} 64 (3): 565--79.
\url{https://doi.org/10.1016/j.molcel.2016.09.027}.

\leavevmode\hypertarget{ref-figliuzzi_micrornas_2013}{}%
Figliuzzi, Matteo, Enzo Marinari, and Andrea De Martino. 2013.
``MicroRNAs as a Selective Channel of Communication Between Competing
RNAs: A Steady-State Theory.'' \emph{Biophysical Journal} 104 (5):
1203--13. \url{https://doi.org/10.1016/j.bpj.2013.01.012}.

\leavevmode\hypertarget{ref-grimson_microrna_2007}{}%
Grimson, Andrew, Kyle Kai-How Farh, Wendy K. Johnston, Philip
Garrett-Engele, Lee P. Lim, and David P. Bartel. 2007. ``MicroRNA
Targeting Specificity in Mammals: Determinants Beyond Seed Pairing.''
\emph{Molecular Cell} 27 (1): 91--105.
\url{https://doi.org/10.1016/j.molcel.2007.06.017}.

\leavevmode\hypertarget{ref-hafner_transcriptome-wide_2010}{}%
Hafner, Markus, Markus Landthaler, Lukas Burger, Mohsen Khorshid, Jean
Hausser, Philipp Berninger, Andrea Rothballer, et al. 2010.
``Transcriptome-Wide Identification of RNA-Binding Protein and MicroRNA
Target Sites by PAR-CLIP.'' \emph{Cell} 141 (1): 129--41.
\url{https://doi.org/10.1016/j.cell.2010.03.009}.

\leavevmode\hypertarget{ref-hausser_analysis_2013}{}%
Hausser, J., A. P. Syed, B. Bilen, and M. Zavolan. 2013. ``Analysis of
CDS-Located miRNA Target Sites Suggests That They Can Effectively
Inhibit Translation.'' \emph{Genome Research} 23 (4): 604--15.
\url{https://doi.org/10.1101/gr.139758.112}.

\leavevmode\hypertarget{ref-helwak_mapping_2013}{}%
Helwak, Aleksandra, Grzegorz Kudla, Tatiana Dudnakova, and David
Tollervey. 2013. ``Mapping the Human miRNA Interactome by CLASH Reveals
Frequent Noncanonical Binding.'' \emph{Cell} 153 (3): 654--65.
\url{https://doi.org/10.1016/j.cell.2013.03.043}.

\leavevmode\hypertarget{ref-lai_understanding_2016}{}%
Lai, Xin, Olaf Wolkenhauer, and Julio Vera. 2016. ``Understanding
microRNA-Mediated Gene Regulatory Networks Through Mathematical
Modelling.'' \emph{Nucleic Acids Research} 44 (13): 6019--35.
\url{https://doi.org/10.1093/nar/gkw550}.

\leavevmode\hypertarget{ref-lewis_conserved_2005}{}%
Lewis, Benjamin P., Christopher B. Burge, and David P. Bartel. 2005.
``Conserved Seed Pairing, Often Flanked by Adenosines, Indicates That
Thousands of Human Genes Are MicroRNA Targets.'' \emph{Cell} 120 (1):
15--20. \url{https://doi.org/10.1016/j.cell.2004.12.035}.

\leavevmode\hypertarget{ref-markus_list_sponge_2017}{}%
Markus List, Marcel Schulz. 2017. ``SPONGE.'' Bioconductor.
\url{https://doi.org/10.18129/B9.bioc.SPONGE}.

\leavevmode\hypertarget{ref-moore_mirnatarget_2015}{}%
Moore, Michael J., Troels K. H. Scheel, Joseph M. Luna, Christopher Y.
Park, John J. Fak, Eiko Nishiuchi, Charles M. Rice, and Robert B.
Darnell. 2015. ``miRNA-Target Chimeras Reveal miRNA 3'-End Pairing as a
Major Determinant of Argonaute Target Specificity.'' \emph{Nature
Communications} 6 (November): 8864.
\url{https://doi.org/10.1038/ncomms9864}.

\leavevmode\hypertarget{ref-nitzan_interactions_2014}{}%
Nitzan, Mor, Avital Steiman-Shimony, Yael Altuvia, Ofer Biham, and Hanah
Margalit. 2014. ``Interactions Between Distant ceRNAs in Regulatory
Networks.'' \emph{Biophysical Journal} 106 (10): 2254--66.
\url{https://doi.org/10.1016/j.bpj.2014.03.040}.

\leavevmode\hypertarget{ref-robinson_modelling_2018}{}%
Robinson, J. M., and W. A. Henderson. 2018. ``Modelling the Structure of
a ceRNA-Theoretical, Bipartite microRNA-mRNA Interaction Network
Regulating Intestinal Epithelial Cellular Pathways Using R
Programming.'' \emph{BMC Research Notes} 11 (1): 19.
\url{https://doi.org/10.1186/s13104-018-3126-y}.

\leavevmode\hypertarget{ref-ruegger_microrna_2012}{}%
Rüegger, Stefan, and Helge Großhans. 2012. ``MicroRNA Turnover: When,
How, and Why.'' \emph{Trends in Biochemical Sciences} 37 (10): 436--46.
\url{https://doi.org/10.1016/j.tibs.2012.07.002}.

\leavevmode\hypertarget{ref-salmena_cerna_2011}{}%
Salmena, Leonardo, Laura Poliseno, Yvonne Tay, Lev Kats, and Pier Paolo
Pandolfi. 2011. ``A ceRNA Hypothesis: The Rosetta Stone of a Hidden RNA
Language?'' \emph{Cell} 146 (3): 353--58.
\url{https://doi.org/10.1016/j.cell.2011.07.014}.

\leavevmode\hypertarget{ref-tay_multilayered_2014}{}%
Tay, Yvonne, John Rinn, and Pier Paolo Pandolfi. 2014. ``The
Multilayered Complexity of ceRNA Crosstalk and Competition.''
\emph{Nature} 505 (7483): 344--52.
\url{https://doi.org/10.1038/nature12986}.

\leavevmode\hypertarget{ref-werfel_preferential_2017}{}%
Werfel, Stanislas, Simon Leierseder, Benjamin Ruprecht, Bernhard Kuster,
and Stefan Engelhardt. 2017. ``Preferential microRNA Targeting Revealed
by in Vivo Competitive Binding and Differential Argonaute
Immunoprecipitation.'' \emph{Nucleic Acids Research} 45 (17): 10218--28.
\url{https://doi.org/10.1093/nar/gkx640}.

\leavevmode\hypertarget{ref-xu_characterization_2014}{}%
Xu, Wenlong, Zixing Wang, and Yin Liu. 2014. ``The Characterization of
microRNA-Mediated Gene Regulation as Impacted by Both Target Site
Location and Seed Match Type.'' Edited by Thomas Preiss. \emph{PLoS ONE}
9 (9): e108260. \url{https://doi.org/10.1371/journal.pone.0108260}.

\end{document}