### 2019-09-10 deneme.Rmd
news to article

---
title: "Network based multifactorial modelling microRNA-target interaction."
author: 
- Selcen Ari^[Department of Bioengineering, Yildiz Technical University, Istanbul, Turkey]
- Alper Yilmaz^[Department of Bioengineering, Yildiz Technical University, Istanbul, Turkey]
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: 
    keep_tex: true
  html_document: default
---

```{r echo= FALSE, results= 'asis'}
install.packages("xtable")
library(xtable)
data("minsamp")

print(xtable::xtable(minsamp, caption="Expression values of elements and interaction factors of miRNA:target interacions", label = "tab:one", 
             table.placement = "l!", 
             caption.placement = "top"), comment= FALSE)
```

\begin{table}[ht]
\centering
\caption{Expression values of elements and interaction factors of miRNA:target interacions} 
\begin{tabular}{rllrrrrr}
  \hline
 & competing & miRNA & Competing\_expression & miRNA\_expression & seed\_type & region & energy \\ 
  \hline
1 & Gene1 & Mir1 & 10000.00 & 1000.00 & 0.43 & 0.30 & -20.00 \\ 
  2 & Gene2 & Mir1 & 10000.00 & 1000.00 & 0.43 & 0.01 & -15.00 \\ 
  3 & Gene3 & Mir1 & 5000.00 & 1000.00 & 0.32 & 0.40 & -14.00 \\ 
  4 & Gene4 & Mir1 & 10000.00 & 1000.00 & 0.23 & 0.50 & -10.00 \\ 
  5 & Gene4 & Mir2 & 10000.00 & 2000.00 & 0.35 & 0.90 & -12.00 \\ 
  6 & Gene5 & Mir2 & 5000.00 & 2000.00 & 0.05 & 0.40 & -11.00 \\ 
  7 & Gene6 & Mir2 & 10000.00 & 2000.00 & 0.01 & 0.80 & -25.00 \\ 
   \hline
\end{tabular}

\end{table}