### 2019-09-22 Supplementary_additional_file.Rmd
additional files

---
title: "Supplementary Experimental Data File"
author: "Selcen Arı"
date: "22 09 2019"
output:
  pdf_document: default
  html_document: default
  word_document: default
bibliography: ref.bib
header-includes: 
  \usepackage{float} \floatplacement{figure}{H} 
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0}     \renewcommand{\thefigure}{S\arabic{figure}}}
  \usepackage{subfig}
---

```{r, include=FALSE, message=FALSE, warning=FALSE}

library(tidygraph)
library(dplyr)
library(tidyr)
library(stringr)
library(readr)
library(readxl)

```

# Arrangement of CLASH dataset [@helwak_mapping_2013]

CLASH dataset was retrieved from [PubMed](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3650559/)

```{r, eval=FALSE}

clashelwak <- read.table("mmc1.txt", comment.char = "#", header = TRUE, skip = 1)

#hg19
```

## Query of Human Genome 19.

```{r,  eval=FALSE}
#HG19
listEnsemblArchives()
listMarts(host = 'http://grch37.ensembl.org')
ensemblgrch37 = useMart(host='http://grch37.ensembl.org', 
                     biomart='ENSEMBL_MART_ENSEMBL',
                    dataset='hsapiens_gene_ensembl')
hg19 <- getBM(attributes = c("ensembl_transcript_id","ensembl_gene_id","chromosome_name", "start_position","end_position", "hgnc_symbol", "entrezgene_id", "strand" ), mart = ensemblgrch37)

```

## Adding miRNA and gene information

```{r,  eval=FALSE}

clashelwak <- clashelwak%>%
 separate(microRNA_name, c("Barcode", "Database", "mirna_name", "type"), sep = "_")%>%
  separate(mRNA_name, c("Ensembl_Gene_Id", "Ensembl_Transcript_Id", "Hugo_Symbol", "mRNA_Type"), sep = "_")

```

MiRNA releases are obtained from [miRBase](http://www.mirbase.org/ftp.shtml). In this step, release 21 (in Human genome 38) is utilised.

```{r, eval=FALSE}
read.table("mirbasehg38.txt", comment.char = "#")%>%
  filter(V3 != "miRNA_primary_transcript")%>%
  separate(V9, c("ID", "Alias", "Name", "Precusor"), sep = ";")%>%
  mutate(ID = substr(ID, 4, length(ID)), Alias = substr(Alias, 7, length(Alias)), Name = substr(Name, 6, length(Name)), Precusor = substr(Precusor, 14, length(Precusor)))%>%
  dplyr::select(chr= V1, start = V4, end = V5, strand = V7, ID, Alias, Name, Precusor)->mirbasehg38

```

CLASH dataset is published in miRBase erlease 15 and Human Genome 19 version.

```{r, eval=FALSE}
read_tsv("mirna_mature.txt", col_names = FALSE)%>%
  filter(startsWith(X2, "hsa"))%>%
  dplyr::select(mirna_ID = X2, mirbase_ID = X3)%>%
  inner_join(mirbasehg38%>% dplyr::select(ID, Name), by= c("mirbase_ID"= "ID"))%>%
  dplyr::select(mirbase_ID, Name)%>%
  distinct()%>%
  inner_join(clashelwak, by = c("mirbase_ID"= "Barcode"))%>%
  dplyr::select(Name, miRNA_seq, Ensembl_Gene_Id, Ensembl_Transcript_Id, Hugo_Symbol, mRNA_seq_extended, chimeras_decompressed, seed_type, seed_basepairs,folding_class, seq_ID, folding_energy, X5.UTR, CDS, X3.UTR)%>%
  inner_join(hg19, by = c("Ensembl_Gene_Id"= "ensembl_gene_id", "Ensembl_Transcript_Id"="ensembl_transcript_id", "Hugo_Symbol"="hgnc_symbol"))%>%
  mutate(region1 = ifelse(X5.UTR == "1", "5UTR", " "), region2= ifelse(X3.UTR == "1", "3UTR", " "),region3= ifelse(CDS =="1", "CDS", " "))%>%
  unite(region, c(region1, region2, region3), sep = "||")%>%
  dplyr::select(chromosome_name, start_position, end_position,strand, Hugo_Symbol, Ensembl_Gene_Id, Ensembl_Transcript_Id, mRNA_seq_extended, Name, miRNA_seq, seq_ID, seed_type,seed_basepairs, folding_class, folding_energy, region)%>%
  as_tibble() -> clashelwakfinal

```

## Converting CLASH data to human genome 38 build.

```{r,  eval=FALSE}
lift19 <- clashelwakfinal%>%
  dplyr::select(1,2,3)%>%
  unite(start_end, c("start_position", "end_position"), sep = "-")%>%
  mutate(Chromosome = paste0("chr",chromosome_name, ""))%>%
  unite(chromosome_name, c("Chromosome","start_end"), sep = ":")

write_tsv(lift19, "lift19.txt")

# Lift over process is made via UCSC liftover tool. (https://genome.ucsc.edu/cgi-bin/hgLiftOver)

lift19_del <- read_tsv("deleted_lift19.txt")
colnames(lift19_del)[1] <- "chromosome_loc"

lift19_del <- lift19_del%>%
  dplyr::filter(startsWith(chromosome_loc, "chr"))%>%
  separate(chromosome_loc, c("Chr", "End"), "-", remove = TRUE)%>%
  separate(Chr, c("Chr", "Start"), ":", remove = TRUE)

lift19_del$Start <- as.numeric(lift19_del$Start)
lift19_del$End <- as.numeric(lift19_del$End)

clashelwakfinal <-clashelwakfinal%>%
  mutate(Chromosome = paste0("chr",chromosome_name, ""))%>%
  dplyr::anti_join(lift19_del, by = c("Chromosome"="Chr", "start_position"="Start", "end_position"="End"))

hg38clash <- read.delim("hg38clashcomp.txt", header = FALSE)

clashelwakfinal <- clashelwakfinal%>% 
  bind_cols(hg38clash)

colnames(clashelwakfinal)[18] <- "HG38build_loc"

clashelwakfinal <- clashelwakfinal%>%
  dplyr::mutate(Genom_build= rep("hg19"))
str(clashelwakfinal)

# Arrangement in dataset

clashelwakfinal%>% dplyr::select(cluster=seq_ID, chromosome = Chromosome, start_position, end_position, strand, hgnc_symbol = Hugo_Symbol, Ensembl_Gene_Id, Ensembl_Transcript_Id, target_seq = mRNA_seq_extended, miRNA= Name, miR_seq =miRNA_seq, seed_type, seed_type2= seed_basepairs, seed_type3= folding_class, Energy = folding_energy, HG38build_loc, Genom_build, region) -> clashelwakfinal

clashelwakfinal$cluster <- as.character(clashelwakfinal$cluster)
clashelwakfinal$strand <- as.character(clashelwakfinal$strand)
clashelwakfinal$target_seq <- as.character(clashelwakfinal$target_seq)
clashelwakfinal$miR_seq <- as.character(clashelwakfinal$miR_seq)
clashelwakfinal$seed_type <- as.character(clashelwakfinal$seed_type)
clashelwakfinal$HG38build_loc <- as.character(clashelwakfinal$HG38build_loc)
clashelwakfinal$seed_type2 <- as.numeric(clashelwakfinal$seed_type2)
clashelwakfinal$seed_type3 <- as.character(clashelwakfinal$seed_type3)
```

## Interpreting the CLASH seed structures in dataset

```{r,  eval=FALSE}
clashelwakfinal%>%
  mutate(seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="I", paste0(seed_type2, "-mer"), seed_type),
         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="II", paste0(seed_type2, "-mer_noncanonical"), seed_type), 
         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="III", paste0(seed_type2, "-mer_noncanonical"), seed_type),
         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="IV", paste0(seed_type2, "-mer_noncanonical"), seed_type),
         seed_type= ifelse (startsWith(seed_type, "no") ,"none", seed_type))%>%
  dplyr::select(-seed_type2, -seed_type3)->clashelwakfinal

```

## Arrangement of CLEAR-CLiP Dataset [@moore_mirnatarget_2015]

CLASH dataset was retrieved from [Nature web page](nature.com/articles/ncomms9864)

```{r,  eval=FALSE}
clearclip <- read_xlsx("CLEAR-CLIP.xlsx")

#Clearclip hg18
```

## Query of Human Genome 18

```{r,  eval=FALSE}
#HG18
listEnsemblArchives()
listMarts(host = 'may2009.archive.ensembl.org')
ensembl54 = useMart(host='may2009.archive.ensembl.org', 
                     biomart='ENSEMBL_MART_ENSEMBL',
                    dataset='hsapiens_gene_ensembl')

hg18 <- getBM(attributes = c("ensembl_transcript_id","ensembl_gene_id","chromosome_name", "start_position","end_position", "hgnc_symbol", "entrezgene", "strand" ), mart = ensembl54)
```

## Adding Genome Information to dataset

```{r,  eval=FALSE}
clearclipfinal <- hg18%>%
  inner_join(clearclip, by= c("entrezgene"= "gene.id", "hgnc_symbol"= "gene.symbol"))%>%
  distinct()
```

## Converting human genome build

```{r,   eval=FALSE}
lift18 <- clearclipfinal%>%
  unite(start_end, c("start_position", "end_position"), sep = "-")%>%
  unite(location, c("chr","start_end"), sep = ":")%>%
  dplyr::select(location)

write_tsv(lift18, "lift18.txt")

deleted_lift18 <- read_tsv("deleted_lift18.txt")

colnames(deleted_lift18)[1] <- "Chromosome_loc"

deleted_lift18 <- deleted_lift18%>%
  dplyr::filter(startsWith(Chromosome_loc, "chr"))%>%
  separate(Chromosome_loc, c("Chr", "End"), "-", remove = TRUE)%>%
  separate(Chr, c("Chr", "Start"), ":", remove = TRUE)

deleted_lift18$Start <- as.numeric(deleted_lift18$Start)
deleted_lift18$End <- as.numeric(deleted_lift18$End)

clearclipfinal <- clearclipfinal%>%
  dplyr::anti_join(deleted_lift18, by = c("chr"="Chr", "start_position"="Start", "end_position"="End"))

hg38clearclip<- read.delim("hg38clearclip.txt", header = FALSE)

clearclipfinal <- clearclipfinal%>% 
  bind_cols(hg38clearclip)

colnames(clearclipfinal)[28] <- "HG38build_loc"

clearclipfinal <- clearclipfinal%>%
  dplyr::mutate(Genom_build= rep("hg18"))

clearclipfinal%>%
  dplyr::select(cluster=cluster.ID, chromosome = chr, start_position, end_position, strand = strand.y, hgnc_symbol, Ensembl_Gene_Id =ensembl_gene_id, Ensembl_Transcript_Id =ensembl_transcript_id, target_seq = target.map, miRNA, miR_seq = miR.map, seed_type = 'seed match', Energy = MFE, HG38build_loc, Genom_build, region) -> clearclipfinal
```

## Seed type manipulation

```{r,  eval=FALSE}
data_frame(seed_type= c("5mer_1","5mer_2", "5mer_3", "6mer", "6mer.indel", "6mer.mm", "6mer_off.mm", "6merA1", "6merA1.indel", "6merA1.mm", "7merA1", "7merA1.indel", "7merA1.mm", "7merm8", "7merm8.indel", "7merm8,mm", "8mer", "8mer.indel", "8mer.mm", "NA" ),
           seed_type_com= c("5-mer", "5-mer_noncanonical", "5-mer_noncanonical", "6-mer", "6-mer_noncanonical", "6-mer_noncanonical", "6-mer_noncanonical", "6-merA1", "6-merA1_noncanonical","6-merA1_noncanonical", "7-merA1", "7-merA1_noncanonical","7-merA1_noncanonical", "7-mer-8m", "7-mer-8m_noncanonical", "7-mer-8m_noncanonical", "8-mer", "8-mer_noncanonical", "8-mer_noncanonical", "none" )) ->clipdata_seed

clearclipfinal%>%
  inner_join(clipdata_seed, by = "seed_type")%>%
  dplyr::select(1:11, seed_type = seed_type_com, Energy, HG38build_loc, Genom_build, region)-> clearclipfinal

clearclipfinal$HG38build_loc <- as.character(clearclipfinal$HG38build_loc)

```

## Integration of two experimental dataset

```{r,  eval=FALSE}

bind_rows(clashelwakfinal, clearclipfinal)%>%distinct() -> experimentalmirnagene

```

## Adding Coefficients of Interaction factors

```{r,  eval=FALSE}
experimentalmirnagene <- experimentalmirnagene%>%
  mutate(region2  = str_replace_all(region, "NA", ""), region3 = str_replace_all(region2, "\\|", ""), region = str_replace_all(region3, c("3'UTR"="3UTR", "5'UTR"="5UTR")))%>%
  dplyr::select(-region2, -region3)%>%
  mutate(region_effect = as.double(ifelse(region %in% c("3UTRCDS", "CDS3UTR", "5UTR3UTR", "CDS5UTR3UTR", "CDS3UTRintron"), "0.93", ifelse(region %in% c("CDS", "CDSintron"), "0.42", ifelse(region %in% c("3UTR","3UTRintron"), "0.84", ifelse(region %in% c("5UTR", "5UTRintron"), "0.01", ifelse(region %in% c("5UTRCDS", "CDS5UTR"), "0.42", ifelse(region %in% c("intron",""), "0.01", ifelse(region %in% c("exon_unclassified", ""), "0.2", NA)))))))))

```

```{r,  eval=FALSE}

seed_type_effect <- data_frame( seed_type = c("5-mer", "5-mer_noncanonical", "6-mer", "6-mer_noncanonical", "6-merA1", "6-merA1_noncanonical", "7-mer", "7-mer_noncanonical", "7-merA1", "7-merA1_noncanonical", "7-mer-8m", "7-mer-8m_noncanonical", "8-mer", "8-mer_noncanonical", "9-mer", "9-mer_noncanonical", "none"),
                                seed_type_effect= c(0.05, 0.04, 0.07, 0.05, 0.07, 0.05, 0.23, 0.19, 0.19, 0.16, 0.25, 0.21, 0.43, 0.35, 0.43,0.35, 0.01)
                                )

experimentalmirnagene%>%
  inner_join(seed_type_effect, by= "seed_type") -> experimentalmirnagene
```

```{r, eval=FALSE}
## Saving dataset
saveRDS(experimentalmirnagene,"data/experimentalmirnagene.RDS")

```

```{r}
readRDS("data/experimentalmirnagene.RDS")->experimentalmirnagene
experimentalmirnagene
```

# REFERENCES