### 2019-12-27 Supplementary_additional_file.Rmd

---
title: "Supplementary Experimental Data File"
author: 
- Selcen Ari Yuka
- Alper Yilmaz 
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
  word_document: default#
bibliography: NAR-sample.bib
header-includes: 
  \usepackage{float} \floatplacement{figure}{H} 
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0}     \renewcommand{\thefigure}{S\arabic{figure}}}
  \usepackage{subfig}
---

```{r, include=FALSE, message=FALSE, warning=FALSE}

library(tidygraph)
library(dplyr)
library(tidyr)
library(stringr)
library(readr)
library(readxl)
library(biomaRt)

```

# Arrangement of CLASH dataset

CLASH dataset was retrieved from [PubMed](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3650559/) [@helwak_mapping_2013].

```{r, eval=FALSE}

clashelwak <- read.table("mmc1.txt", comment.char = "#", header = TRUE, 
                         skip = 1, stringsAsFactors = FALSE)

#hg19
```

## Query of Human Genome 19.

Human genome 19 information was handled through biomaRt package.

```{r,  eval=FALSE}
#HG19
listEnsemblArchives()
listMarts(host = 'http://grch37.ensembl.org')
ensemblgrch37 = useMart(host='http://grch37.ensembl.org', 
                     biomart='ENSEMBL_MART_ENSEMBL',
                    dataset='hsapiens_gene_ensembl')
hg19 <- getBM(attributes = c("ensembl_transcript_id","ensembl_gene_id",
                             "chromosome_name", "start_position","end_position", 
                             "hgnc_symbol", "entrezgene_id", "strand" ), mart = ensemblgrch37)

```

## Adding miRNA and gene information

```{r,  eval=FALSE}

clashelwak <- clashelwak%>%
 separate(microRNA_name, c("Barcode", "Database", 
                           "mirna_name", "type"), sep = "_")%>%
  separate(mRNA_name, c("Ensembl_Gene_Id", "Ensembl_Transcript_Id", 
                        "Hugo_Symbol", "mRNA_Type"), sep = "_")

```

MiRNA releases are obtained from [miRBase](http://www.mirbase.org/ftp.shtml). In this step, release 21 (in Human genome 38) was downloaded.

```{r, eval=FALSE}
read.table("mirbasehg38.txt", comment.char = "#")%>%
  filter(V3 != "miRNA_primary_transcript")%>%
  separate(V9, c("ID", "Alias", "Name", "Precusor"), sep = ";")%>%
  mutate(ID = substr(ID, 4, length(ID)), 
         Alias = substr(Alias, 7, length(Alias)), 
         Name = substr(Name, 6, length(Name)), 
         Precusor = substr(Precusor, 14, length(Precusor)))%>%
  dplyr::select(chr= V1, start = V4, end = V5, 
                strand = V7, ID, Alias, Name, Precusor)->mirbasehg38

```

CLASH dataset is published in miRBase release 15 and Human Genome 19 version.

```{r, eval=FALSE}
read_tsv("mirna_mature.txt", col_names = FALSE)%>%
  filter(startsWith(X2, "hsa"))%>%
  dplyr::select(mirna_ID = X2, mirbase_ID = X3)%>%
  inner_join(mirbasehg38%>% dplyr::select(ID, Name), 
             by= c("mirbase_ID"= "ID"))%>%
  dplyr::select(mirbase_ID, Name)%>%
  distinct()%>%
  inner_join(clashelwak, by = c("mirbase_ID"= "Barcode"))%>%
  dplyr::select(Name, miRNA_seq, Ensembl_Gene_Id, 
                Ensembl_Transcript_Id, Hugo_Symbol, 
                mRNA_seq_extended, chimeras_decompressed, 
                seed_type, seed_basepairs,folding_class, 
                seq_ID, folding_energy, 
                X5.UTR, CDS, X3.UTR)%>%
  inner_join(hg19, by = c("Ensembl_Gene_Id"= "ensembl_gene_id", 
                          "Ensembl_Transcript_Id"="ensembl_transcript_id", 
                          "Hugo_Symbol"="hgnc_symbol"))%>%
  mutate(region1 = ifelse(X5.UTR == "1", "5UTR", " "), 
         region2= ifelse(X3.UTR == "1", "3UTR", " "),
         region3= ifelse(CDS =="1", "CDS", " "))%>%
  unite(region, c(region1, region2, region3), sep = "||")%>%
  dplyr::select(chromosome_name, start_position, 
                end_position, strand, Hugo_Symbol, 
                Ensembl_Gene_Id, Ensembl_Transcript_Id, 
                mRNA_seq_extended, Name, miRNA_seq, 
                seq_ID, seed_type,seed_basepairs, 
                folding_class, folding_energy, region)%>%
  as_tibble() -> clashelwakfinal

```

## Converting CLASH data to human genome 38 build.

There are different liftover methods for conversion among Human Genome builds. We preffered to use [UCSC liftover tool](https://genome.ucsc.edu/cgi-bin/hgLiftOver)

```{r,  eval=FALSE}

#Obtaining chromosomal locations from miRNA:target interaction dataset.

lift19 <- clashelwakfinal%>%
  dplyr::select(1,2,3)%>%
  unite(start_end, c("start_position", "end_position"), sep = "-")%>%
  mutate(Chromosome = paste0("chr",chromosome_name, ""))%>%
  unite(chromosome_name, c("Chromosome","start_end"), sep = ":")

write_tsv(lift19, "lift19.txt")

#After we searched this file in UCSC browser, the output loaded (lift19_del deleted regions on the HG38 genome build; hg38clashcomp new locations of the genes on HG38)

lift19_del <- read_tsv("deleted_lift19.txt")
colnames(lift19_del)[1] <- "chromosome_loc"

lift19_del <- lift19_del%>%
  dplyr::filter(startsWith(chromosome_loc, "chr"))%>%
  separate(chromosome_loc, c("Chr", "End"), "-", remove = TRUE)%>%
  separate(Chr, c("Chr", "Start"), ":", remove = TRUE)

lift19_del$Start <- as.numeric(lift19_del$Start)
lift19_del$End <- as.numeric(lift19_del$End)

#removing deleted location from CLASH dataset

clashelwakfinal <-clashelwakfinal%>%
  mutate(Chromosome = paste0("chr",chromosome_name, ""))%>%
  dplyr::anti_join(lift19_del, by = c("Chromosome"="Chr", 
                                      "start_position"="Start", 
                                      "end_position"="End"))

hg38clash <- read.delim("hg38clashcomp.txt", header = FALSE, stringsAsFactors = FALSE)

# adding new location information:

clashelwakfinal <- clashelwakfinal%>% 
  bind_cols(hg38clash)

colnames(clashelwakfinal)[18] <- "HG38build_loc"

clashelwakfinal <- clashelwakfinal%>%
  dplyr::mutate(Genom_build= rep("hg19"))

# Arrangement in dataset

clashelwakfinal%>% dplyr::select(cluster=seq_ID, chromosome = Chromosome, 
                                 start_position, end_position, 
                                 strand, hgnc_symbol = Hugo_Symbol, 
                                 Ensembl_Gene_Id, Ensembl_Transcript_Id, 
                                 target_seq = mRNA_seq_extended, miRNA= Name, 
                                 miR_seq =miRNA_seq, seed_type, 
                                 seed_type2= seed_basepairs, 
                                 seed_type3= folding_class, 
                                 Energy = folding_energy, 
                                 HG38build_loc, 
                                 Genom_build, region) -> clashelwakfinal

clashelwakfinal$strand <- as.character(clashelwakfinal$strand)

str(clashelwakfinal)
```

## Interpreting the CLASH seed structures in dataset

```{r,  eval=FALSE}
clashelwakfinal%>%
  mutate(seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="I", 
                            paste0(seed_type2, "-mer"), seed_type),
         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="II", 
                            paste0(seed_type2, "-mer_noncanonical"), seed_type), 
         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="III", 
                            paste0(seed_type2, "-mer_noncanonical"), seed_type),
         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="IV", 
                            paste0(seed_type2, "-mer_noncanonical"), seed_type),
         seed_type= ifelse (startsWith(seed_type, "no") ,"none", seed_type))%>%
  dplyr::select(-seed_type2, -seed_type3)->clashelwakfinal

```

## Arrangement of CLEAR-CLiP Dataset [@moore_mirnatarget_2015]

CLASH dataset was retrieved from [Nature article](nature.com/articles/ncomms9864)

```{r,  eval=FALSE}
clearclip <- read_xlsx("CLEAR-CLIP.xlsx")

#Clearclip hg18
```

## Query of Human Genome 18

```{r,  eval=FALSE}
#HG18
listEnsemblArchives()
listMarts(host = 'may2009.archive.ensembl.org')
ensembl54 = useMart(host='may2009.archive.ensembl.org', 
                     biomart='ENSEMBL_MART_ENSEMBL',
                    dataset='hsapiens_gene_ensembl')

hg18 <- getBM(attributes = c("ensembl_transcript_id","ensembl_gene_id",
                             "chromosome_name", "start_position","end_position", 
                             "hgnc_symbol", "entrezgene", "strand" ), mart = ensembl54)
```

## Adding Genome Information to dataset

```{r,  eval=FALSE}
clearclipfinal <- hg18%>%
  inner_join(clearclip, by= c("entrezgene"= "gene.id", 
                              "hgnc_symbol"= "gene.symbol"))%>%
  distinct()
```

## Converting human genome build

```{r,   eval=FALSE}

#Obtaining chromosomal locations from miRNA:target interaction dataset.

lift18 <- clearclipfinal%>%
  unite(start_end, c("start_position", "end_position"), sep = "-")%>%
  unite(location, c("chr","start_end"), sep = ":")%>%
  dplyr::select(location)

write_tsv(lift18, "lift18.txt")

#After we searched this file in UCSC browser, the output loaded (deleted_lift18 deleted regions on the HG38 genome build; hg38clearclip new locations of the genes on HG38)

deleted_lift18 <- read_tsv("deleted_lift18.txt")

colnames(deleted_lift18)[1] <- "Chromosome_loc"

deleted_lift18 <- deleted_lift18%>%
  dplyr::filter(startsWith(Chromosome_loc, "chr"))%>%
  separate(Chromosome_loc, c("Chr", "End"), "-", remove = TRUE)%>%
  separate(Chr, c("Chr", "Start"), ":", remove = TRUE)

deleted_lift18$Start <- as.numeric(deleted_lift18$Start)
deleted_lift18$End <- as.numeric(deleted_lift18$End)

#removing deleted location from CLEAR-CLiP dataset

clearclipfinal <- clearclipfinal%>%
  dplyr::anti_join(deleted_lift18, by = c("chr"="Chr", 
                                          "start_position"="Start", 
                                          "end_position"="End"))

hg38clearclip<- read.delim("hg38clearclip.txt", header = FALSE, 
                           stringsAsFactors = FALSE)

clearclipfinal <- clearclipfinal%>% 
  bind_cols(hg38clearclip)

colnames(clearclipfinal)[28] <- "HG38build_loc"

# adding new location information:

clearclipfinal <- clearclipfinal%>%
  dplyr::mutate(Genom_build= rep("hg18"))

# Arrangement in dataset

clearclipfinal%>%
  dplyr::select(cluster=cluster.ID, chromosome = chr, start_position, 
                end_position, strand = strand.y, hgnc_symbol, Ensembl_Gene_Id =ensembl_gene_id, 
                Ensembl_Transcript_Id =ensembl_transcript_id, target_seq = target.map, 
                miRNA, miR_seq = miR.map, seed_type = 'seed match', Energy = MFE, 
                HG38build_loc, Genom_build, region) -> clearclipfinal

```

## Seed type manipulation in CLEAR-CLiP dataset

In CLEAR-CLiP dataset, seed types were shown in detail. We adjusted as canonical and non-canonical.

```{r,  eval=FALSE}
data_frame(seed_type= c("5mer_1","5mer_2", "5mer_3", "6mer", "6mer.indel", "6mer.mm", 
                        "6mer_off.mm", "6merA1", "6merA1.indel", "6merA1.mm", "7merA1", 
                        "7merA1.indel", "7merA1.mm", "7merm8", "7merm8.indel", "7merm8,mm", 
                        "8mer", "8mer.indel", "8mer.mm", "NA" ),
           seed_type_com= c("5-mer", "5-mer_noncanonical", "5-mer_noncanonical", "6-mer", 
                            "6-mer_noncanonical", "6-mer_noncanonical", "6-mer_noncanonical", 
                            "6-merA1", "6-merA1_noncanonical","6-merA1_noncanonical", "7-merA1",
                            "7-merA1_noncanonical","7-merA1_noncanonical", "7-mer-8m", 
                            "7-mer-8m_noncanonical", "7-mer-8m_noncanonical", "8-mer", 
                            "8-mer_noncanonical", "8-mer_noncanonical", "none" )) ->clipdata_seed

clearclipfinal%>%
  inner_join(clipdata_seed, by = "seed_type")%>%
  dplyr::select(1:11, seed_type = seed_type_com, Energy, 
                HG38build_loc, Genom_build, region)-> clearclipfinal

```

## Integration of two experimental dataset

```{r,  eval=FALSE}

bind_rows(clashelwakfinal, clearclipfinal)%>%distinct() -> experimentalmirnagene

```

## Adding Coefficients of Interaction factors

Energy values in miRNA:target pairs are represented by high-throughput studies [@helwak_mapping_2013; @moore_mirnatarget_2015] which are utilized in this study. On the other hand, we have specified the other interaction factors, seed type and location of binding region on the target, as numeric values based on the previous studies.Grimson et al. [-@grimson_microrna_2007] have compared the effect of seed types on target repression with few miRNA had canonical seed pairing in their study. Additionally, Bartel et al. [-@bartel_micrornas:_2009] and Betel et al. [-@betel2010comprehensive] have studied on functional and non-functional seed interactions. Based on results of these studies we have arranged seed types of miRNA:target interactions as numeric values. We also have redefined location of binding region on the target as numeric values, based on studies of Hausser et al. [-@hausser_analysis_2013] and Helwak et al. [-@helwak_mapping_2013]. With this process, we have handled this integrated dataset in context of competitor behaviors and functionality of interactions.

In this step we added numeric intraction values at followings

Fistly, we organized these values due to the fact that the regions were defined differently in two datasets. After that, region effect was added as numeric values (shown in Table S3).

```{r,  eval=FALSE}
experimentalmirnagene <- experimentalmirnagene%>%
  mutate(region2  = str_replace_all(region, "NA", ""), 
         region3 = str_replace_all(region2, "\\|", ""), 
         region = str_replace_all(region3, c("3'UTR"="3UTR", "5'UTR"="5UTR")))%>%
  dplyr::select(-region2, -region3)%>%
  mutate(region_effect = as.double(ifelse(region %in% c("3UTRCDS", "CDS3UTR", "5UTR3UTR", "CDS5UTR3UTR", "CDS3UTRintron"), 
                                  "0.93", ifelse(region %in% c("CDS", "CDSintron"), 
                                  "0.42", ifelse(region %in% c("3UTR","3UTRintron"), 
                                  "0.84", ifelse(region %in% c("5UTR", "5UTRintron"), 
                                  "0.01", ifelse(region %in% c("5UTRCDS", "CDS5UTR"),
                                  "0.42", ifelse(region %in% c("intron",""), 
                                  "0.01", ifelse(region %in% c("exon_unclassified", ""), "0.2", NA)))))))))

```

Secondly, we organized seed type interactions in *Seed type manipulation* section for CLEAR-CLiP dataset to show as found in CLASH dataset. Same type formatted values added dataset as numeric values (shown in Table S2).

```{r,  eval=FALSE}

seed_type_effect <- data_frame( seed_type = c("5-mer", "5-mer_noncanonical", 
                                              "6-mer", "6-mer_noncanonical", 
                                              "6-merA1", "6-merA1_noncanonical", 
                                              "7-mer", "7-mer_noncanonical", "7-merA1", 
                                              "7-merA1_noncanonical", "7-mer-8m", 
                                              "7-mer-8m_noncanonical", "8-mer", 
                                              "8-mer_noncanonical", "9-mer", 
                                              "9-mer_noncanonical", "none"),
                                seed_type_effect= c(0.05, 0.04, 0.07, 0.05, 
                                                    0.07, 0.05, 0.23, 0.19, 
                                                    0.19, 0.16, 0.25, 0.21, 
                                                    0.43, 0.35, 0.43,0.35, 0.01)
                                )

experimentalmirnagene%>%
  inner_join(seed_type_effect, by= "seed_type") -> experimentalmirnagene
```

```{r, eval=FALSE}
## Saving dataset
saveRDS(experimentalmirnagene,"data/experimentalmirnagene.RDS")

```

```{r}
readRDS("data/experimentalmirnagene.RDS")->experimentalmirnagene
experimentalmirnagene
```

The context of dataset is shown in Table S5 in Supplementary Tables.

# REFERENCES