### 2019-12-27 Supplementary_figures.Rmd

---
title: "Supplementary Figures"
author: 
- Selcen Ari Yuka
- Alper Yilmaz 
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
  word_document: default
header-includes: 
  \usepackage{float} \floatplacement{figure}{H} 
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0}     \renewcommand{\thefigure}{S\arabic{figure}}}
  \usepackage{subfig}
---

\beginsupplement

![Workflow for simulation of competing endogenous RNA regulations. Graph object in steps 2-6 is saved and updated continuously.](workflow1.png){#figs1}

```{r, include=FALSE, message=FALSE, warning=FALSE}

library(tidygraph)
library(dplyr)
library(tidyr)
library(stringr)

```

```{r, message=FALSE, warning=FALSE}
#install.packages("devtools")
#devtools::install_github("selcenari/ceRNAnetsim")
library(ceRNAnetsim)
```

## *minsamp* dataset analysis in absence of interaction factors.

```{r}
minsamp<-readRDS("data/minsamp.RDS")
```

```{r, message=FALSE, warning=FALSE, fig.cap= "Minimal Dataset in Steady-state", fig.width=5, fig.height=4}
priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression)%>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4",
            title = "Minimal dataset in\nsteady-state conditions")
```

```{r, message=FALSE, warning=FALSE, fig.keep='all', fig.cap= "Gene2 Upregulation on Minimal Dataset", fig.width=5, fig.height=4}
priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression)%>%
  update_how("Gene2", 2)%>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4", 
            Upregulation = "red", 
            title = "Gene2 Upregulation\nwithout interaction factors")
```

```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2,  fig.show='hold', fig.cap = "Sequential iteration of minsamp data. A) First response of system to Gene2 upregulation (2nd iteration). B) Spreading of perturbation on system (3th iteration)", fig.subcap= NULL}

priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression) %>%
  update_how("Gene2", 2) %>%
  simulate(cycle=1) %>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4", 
            Upregulation = "red", 
            Downregulation = "blue", 
            title = "A")

priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression)%>%
  update_how("Gene2", 2)%>%
  simulate(2)%>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4", 
            Upregulation = "red", 
            Downregulation = "blue", 
            title = "B")

```

## *minsamp* dataset analysis with interaction factors.

```{r, message=FALSE, warning=FALSE, fig.cap= "Minimal Dataset with interaction factors in Steady-state", fig.width=5, fig.height=4}
priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression, 
              aff_factor = c(energy, seed_type), 
              deg_factor = region)%>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4", 
            title = "Minimal dataset with interaction\nfactors in steady-state conditions")

```

```{r, message=FALSE, warning=FALSE, fig.cap= " When Gene2 is upregulated on Minimal Dataset with interaction factors", fig.width=5, fig.height=4}
priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression, 
              aff_factor = c(energy, seed_type), 
              deg_factor = region)%>%
  update_how("Gene2", 2)%>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4", 
            Upregulation = "red", 
            title = "Gene2 Upregulation\nwith interaction factors")
```

```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.show='hold', fig.cap = "Sequential iteration of minsamp data with interaction factors A)First response of system to Gene2 upregulation (2nd iteration). B)Spreading of perturbation on system (3th iteration)"}
priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression, 
              aff_factor = c(energy, seed_type), 
              deg_factor = region)%>%
  update_how("Gene2", 2)%>%
  simulate(cycle = 1)%>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4", 
            Upregulation = "red", 
            Downregulation = "blue", 
            title = "A")

priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression, 
              aff_factor = c(energy, seed_type), 
              deg_factor = region)%>%
  update_how("Gene2", 2)%>%
  simulate(cycle = 2)%>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4", 
            Upregulation = "red", 
            Downregulation = "blue", 
            title = "B")

```

## Common target perturbation in *minsamp* dataset.

```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.show='hold', fig.cap = "Perturbation of Gene4 on minsamp data with interaction factors. A) Network at steady-state. B) Upregulation of Gene4. C) Primary response of network to upregulation of Gene4. D) Re-regulation of whole nodes on system (3th iteration)"}

priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression, 
              aff_factor = c(energy, seed_type), 
              deg_factor = region)%>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4", 
            Upregulation = "red", 
            title = "A")

priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression, 
              aff_factor = c(energy, seed_type), 
              deg_factor = region)%>%
  update_how("Gene4", 2)%>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4", 
            Upregulation = "red", 
            title = "B")

priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression, 
              aff_factor = c(energy, seed_type), 
              deg_factor = region)%>%
  update_how("Gene4", 2)%>%
  simulate(cycle=1)%>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4", 
            Upregulation = "red", 
            title = "C")

priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression, 
              aff_factor = c(energy, seed_type), 
              deg_factor = region)%>%
  update_how("Gene4", 2)%>%
  simulate(cycle=2)%>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4", 
            Upregulation = "red", 
            title = "D")
```

## Determination of optimum iteration for Upregulation of SERPINE2 gene from Breast cancer patient dataset

```{r}
E9GE_mirnagenenormal<-readRDS("data/E9GE_mirnagenenormal.RDS")
head(E9GE_mirnagenenormal)
```

```{r, fig.cap= "Percentage of affected nodes of each iteration for SERPINE2 Gene"}
as.data.frame(E9GE_mirnagenenormal)%>%
 priming_graph(competing_count = GE_normal, 
               miRNA_count = mirna_RPM, 
               aff_factor = c(Energy, seed_type_effect), 
               deg_factor = region_effect)%>%
 update_how("SERPINE2",2.75) %>%
 simulate(150) %>%
 find_iteration(limit=1, plot= TRUE) 
#limit=0 describes the change that is not taken into account.
```

## Highly effective perturbing elements in network

![Interactions of genes that have high perturbation efficiency. Critical gene, represents crucial genes for breast cancer,provided by eDGAR; TFs, transcription factors;  Regulated by TFs, regulated genes with these transcription factors; Physical interaction with critical genes; shared genes in physical interactions between critical genes ](highly_effective_genes.png){#figs10}