### 2020-10-26 figure_manuscript/codes_of_figures.Rmd
Reproducible rmd files and datasets for figs

---
title: "Main Figures"
author:
- Selcen Ari
- Alper Yilmaz
date: "`r Sys.Date()`"
output: pdf_document
---

# Figures

This file consists analysis steps and codes of figures in manuscript.

```{r, include=FALSE, message=FALSE, warning=FALSE}
library(janitor)
library(tidyverse)
library(purrr)
library(readr)
library(biomaRt)
library(TCGAbiolinks)
library(ceRNAnetsim)
library(dplyr)
library(ggVennDiagram)
library(ggraph)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
#knitr::opts_chunk$set(fig.pos = 'H')
knitr::opts_chunk$set(cache=FALSE)
```

```{r}

mirtarbase_node_perturbations <- readRDS("mirtarbase_node_perturbations.RDS") # data includes results of find_node_perturbation() function for tumor and normal samples of each patient.

```

## Venn diagrams **Figure 4A**

```{r}
mirtarbase_node_perturbations%>%
  group_by(name)%>%
  summarise(perturbed = sum(perturbed_count >0, na.rm = TRUE))%>%
  ungroup()->gene_dist

mirtarbase_node_perturbations%>%
  left_join(gene_dist, by ="name")%>%
  filter(perturbed> 10, perturbed_count > 78)%>%
  distinct(name)->effective_nodes
```

+ Detecting highly perturbing gene nodes in tumor tissues

```{r}
mirtarbase_node_perturbations%>%
  left_join(gene_dist, by = "name")%>%
  mutate(tissue_type = ifelse(endsWith(file_name, "01A"), "Tumor", "Normal"),
         node_type = ifelse(name %in% effective_nodes$name, "effective", "non_effective"))%>%
  filter(node_type == "effective", perturbed_count > 78, perturbed > 1)%>% 
  filter(tissue_type != "Normal", startsWith(name, "ENS"))%>%
  distinct(name)%>%
  pull()->lim_1_gene_tumor
```
+ Detecting highly perturbing gene nodes in normal tissues

```{r}
mirtarbase_node_perturbations%>%
  left_join(gene_dist, by = "name")%>%
  mutate(tissue_type = ifelse(endsWith(file_name, "01A"), "Tumor", "Normal"),
         node_type = ifelse(name %in% effective_nodes$name, "effective", "non_effective"))%>%
  filter(node_type == "effective", perturbed_count > 78, perturbed > 1)%>% 
  filter(tissue_type == "Normal", startsWith(name, "ENS"))%>%
  distinct(name)%>%
  pull()->lim_1_gene_normal
```

### Venn diagram of genes

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width=8, fig.height=8}
x_gene = list(" "= lim_1_gene_tumor, " " = lim_1_gene_normal)

pa1 <- ggVennDiagram(x_gene, label = "count", category.names = c("Tumor","Normal"))+
  scale_fill_gradient(low='bisque1',high = 'lightsalmon')+
  theme(legend.position="none", text = element_text(""), plot.margin = margin(0, 0, 0, 0, "mm"))

pa1

```

+ Detecting highly perturbing mirna nodes in tumor tissues

```{r}

mirtarbase_node_perturbations%>%
  left_join(gene_dist, by = "name")%>%
  mutate(tissue_type = ifelse(endsWith(file_name, "01A"), "Tumor", "Normal"),
         node_type = ifelse(name %in% effective_nodes$name, "effective", "non_effective"))%>%
  filter(node_type == "effective", perturbed_count > 78, perturbed > 1)%>% 
  filter(tissue_type != "Normal", startsWith(name, "hsa"))%>%
  distinct(name)%>%
  pull()->lim_1_mirna_tumor

```

+ Detecting highly perturbing mirna nodes in normal tissues

```{r}

mirtarbase_node_perturbations%>%
  left_join(gene_dist, by = "name")%>%
  mutate(tissue_type = ifelse(endsWith(file_name, "01A"), "Tumor", "Normal"),
         node_type = ifelse(name %in% effective_nodes$name, "effective", "non_effective"))%>%
  filter(node_type == "effective", perturbed_count > 78, perturbed > 1)%>% 
  filter(tissue_type == "Normal", startsWith(name, "hsa"))%>%
  distinct(name)%>%
  pull()->lim_1_mirna_normal

```

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width=8, fig.height=8}
x_mirna = list(" "= lim_1_mirna_tumor, " " = lim_1_mirna_normal)

pa2 <- ggVennDiagram(x_mirna, label = "count", category.names = c("Tumor","Normal"))+
  scale_fill_gradient(low='lightskyblue1',high = 'lightskyblue3')+
  theme(legend.position="none", text = element_text(""), plot.margin = margin(0, 0, 0, 0, "cm"))

pa2
```
## Functional Annotation Analysis

```{r}
Anno_tumor_specific <- readRDS("Anno_tumor_specific.RDS") # includes functional annotation of 29 genes found as tumor-specific.
```

#Fuctional annotation of Tumor specific 29 genes. **Figure 4B**

## Tumor only:

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=10,fig.show='hold'}
Anno_tumor_specific%>%
  filter(startsWith(Category, c("KEGG", "GO")), !str_detect(Category, "CC_DIRECT"))%>%
  filter(!str_detect(Category, "MF_DIRECT"))%>%
  clean_names()%>%
  group_by(category)%>%
  filter(row_number()<6)->tumor_only_annotation

position <- rev(c("hsa05219:Bladder cancer", "hsa05212:Pancreatic cancer", "hsa04510:Focal adhesion", "hsa05206:MicroRNAs in cancer", "hsa05200:Pathways in cancer", "GO:0071230~cellular response to amino acid stimulus", 	"GO:0043154~negative regulation of cysteine-type endopeptidase activity involved in apoptotic process", "GO:0001701~in utero embryonic development", 	"GO:0043066~negative regulation of apoptotic process", "GO:0000122~negative regulation of transcription from RNA polymerase II promoter"))

tumor_only_annotation%>%
  group_by(category)%>%
  arrange(-desc(fold_enrichment))

tumor_only_annotation%>%
  ggplot(aes(x = fold_enrichment, y = term))+
  geom_col(aes(color = category, fill= category))+
  geom_text(aes(label=round(p_value, 5)), hjust = 0.9, color ="darkslategrey", size =3)+
  xlab("Fold Enrichment")+
  ylab("Annotation Term")+
  theme(panel.grid = element_blank(), axis.title = element_text(size =20), axis.text = element_text(size = 12),legend.position = "none")+
  scale_y_discrete(labels=scales::wrap_format(40),position = "left", limits = position)

```

## Network construction with overall functional annotation: **Figure 3**

```{r}

significant_node_graph <- readRDS("significant_node_graph.RDS") #consists network of highly perturbing nodes.
hg19 <- readRDS("hg19.RDS") #For id matching (obtained via bioMaRt package)

```

## Annotation network:

```{r}

Annotation_overall <- readRDS("Annotation_overall.RDS") # consists functional annotation of 145 high perturbing nodes.

Annotation_overall%>%
  filter(startsWith(Category, c("KEGG", "GO")), !str_detect(Category, "CC_DIRECT"))%>%
  filter(!str_detect(Category, "MF_DIRECT"))%>%
  clean_names()%>%
  group_by(category)%>%
  filter(row_number()<6)%>%
  dplyr::select(term, Gene_Count = count, p_value, genes)%>%
  mutate(genes2 = str_split(genes, ", "))%>%
  unnest()%>%
  ungroup%>%
  mutate(genes2 = str_trim(genes2, side = "both"), 
         genes2 = ifelse(endsWith(genes2, ","), substr(genes2, 0, (nchar(genes2)-1)), genes2),
         interaction_type = term, 
         Ensembl_Gene_Id = genes2, 
         competing_name = genes2)%>%
  dplyr::select(interaction= term, Ensembl_Gene_Id, interaction_type, competing_name)->gene_top_annotation_graph

```

```{r, warning=FALSE}

significant_node_graph%>%
  bind_rows(gene_top_annotation_graph)%>%
  as_tbl_graph()%>%
  mutate(type = ifelse(startsWith(name, "hsa"), "KEGG", "GO"),
         type = ifelse(startsWith(name, "hsa-"), "miRNA", type),
         type = ifelse(startsWith(name, "ENSG"), "Gene", type))%>%
  mutate(centrality= centrality_degree(mode = "all"),
         annotation = ifelse(name %in% gene_top_annotation_graph$Ensembl_Gene_Id, "top_annotation", "other"))%>%
  as_tibble()%>%
  filter(annotation == "top_annotation", centrality >10)%>% 
  dplyr::select(name)%>%
  pull()->top_important

significant_node_graph%>%
  bind_rows(gene_top_annotation_graph)%>%
  as_tbl_graph()%>%
  mutate(type = ifelse(startsWith(name, "hsa"), "KEGG", "GO"),
         type = ifelse(startsWith(name, "hsa-"), "miRNA", type),
         type = ifelse(startsWith(name, "ENSG"), "Gene", type))%>%
  mutate(centrality= centrality_degree(mode = "all"),
         annotation = ifelse(name %in% gene_top_annotation_graph$Ensembl_Gene_Id, "Enriched Genes", "other"))%>%
  left_join(hg19, by =c("name"="ensembl_gene_id"))%>%
  filter( type %in% c("miRNA", "Gene"))%>%
  ggraph(layout = "kk")+ 
  geom_edge_link(colour = "seashell2")+
  geom_node_point(color = "seashell2")+
  geom_node_point(aes(filter= annotation == "Enriched Genes", size = centrality, color = "Enriched Genes"), shape = 16)+
  geom_node_point(aes(filter= type == "miRNA", color = "miRNA"), shape = 17, size = 2)+
  geom_node_point(aes(filter= type == "GO", size = centrality, color = "GO"), shape = 18)+
  geom_node_point(aes(filter= type == "KEGG", size = centrality, color = "KEGG"), shape = 15)+
  theme_graph(base_family = "sans")+
  geom_node_text(aes(filter= name %in% top_important, label = hgnc_symbol), size =3.5, repel = TRUE)+
  theme(plot.margin = margin(0, 0, 0, 0, "cm"), legend.position = "none")+
  guides(size=FALSE, shape = FALSE)+
  scale_color_manual(name = "Node types", values = c("Enriched Genes" = "sandybrown", "miRNA"= "dodgerblue2", "GO"= "green", "KEGG"= "red"))

```

# Heat-maps for miRNAs

+Detecting perturbing miRNAs

```{r, warning=FALSE}
mirtarbase_node_perturbations%>%
  left_join(gene_dist, by = "name")%>%
  mutate(tissue_type = ifelse(endsWith(file_name, "01A"), "Tumor", "Normal"),
         node_type = ifelse(name %in% effective_nodes$name, "effective", "non_effective"))%>%
  filter(node_type == "effective", perturbed_count > 78, perturbed > 1)%>%
  filter(tissue_type != "Normal", startsWith(name, "hsa"))%>%
  dplyr::select(file_name, name,tissue_type, perturbed_count)->mirna_tumor

mirtarbase_node_perturbations%>%
  left_join(gene_dist, by = "name")%>%
  mutate(tissue_type = ifelse(endsWith(file_name, "01A"), "Tumor", "Normal"),
         node_type = ifelse(name %in% effective_nodes$name, "effective", "non_effective"))%>%
  filter(node_type == "effective", perturbed_count > 78, perturbed > 1)%>%
  filter(tissue_type == "Normal", startsWith(name, "hsa"))%>%
  dplyr::select(file_name, name,tissue_type, perturbed_count)->mirna_normal

```

## Heat map Figure 4C

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=9,fig.show='hold'}
mirna_tumor%>%
  full_join(mirna_normal, by =c("name", "file_name"), suffix= c("_tumor", "_normal"), fill.na = 0)%>%
  dplyr::select(-3, -5)%>%
  mutate(perturbed_count_tumor = ifelse(is.na(perturbed_count_tumor), 1, perturbed_count_tumor),
         perturbed_count_normal = ifelse(is.na(perturbed_count_normal), 1, perturbed_count_normal),
         log_FC= log2(perturbed_count_tumor/perturbed_count_normal))%>%
  ggplot(aes(x = file_name, y = name, fill = log_FC))+
  geom_tile(aes(colour = log_FC, fill = log_FC))+coord_fixed(ratio = 1)+
  theme_test()+
  theme(axis.text.x = element_blank(), #element_text(angle = 90, vjust = 0, hjust=0)
        plot.title = element_text(hjust = 0.5, size =20), 
        axis.title = element_text(size =20))+
  scale_colour_gradientn(colours = c("green","white", "red"), aesthetics = c("colour", "fill"), na.value = "grey")+
  ylab("Highly pertubing miRNA nodes")+
  xlab("Patients")+
  theme( legend.title = element_blank(), panel.background = element_rect(fill = "white"), plot.margin = margin(0, 0, 0, 0, "cm"))

```

## Heat map Figure 4D

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=10, fig.show='hold'}

mirna_tumor%>%
  full_join(mirna_normal, by =c("name", "file_name"), suffix= c("_tumor", "_normal"), fill.na = 0)%>%
  dplyr::select(-3, -5)%>%
  pivot_longer(cols = c("perturbed_count_tumor", "perturbed_count_normal"), names_to = "situ", values_to = "perturbed_count")%>%
  mutate(perturbed_count_log = log2(perturbed_count),
         situ = str_replace(situ, "perturbed_count_normal", 'Perturbed count in normal'),
         situ = str_replace(situ, "perturbed_count_tumor", 'Perturbed count in tumor'),
         perturbed_count = as.numeric(str_replace_na(perturbed_count, 0)),
         perturbed_count_log = as.numeric(str_replace_na(perturbed_count_log, 0)))%>%
  ggplot(aes(x = file_name, y = name, fill = perturbed_count_log))+
  geom_tile(aes(colour = perturbed_count_log, fill = perturbed_count_log))+coord_fixed(ratio = 1)+
  theme_test()+
  theme(axis.text.x = element_blank(), #element_text(angle = 90, vjust = 0, hjust=0)
        plot.title = element_text(hjust = 0.5, size =20), 
        axis.title = element_text(size =20))+
  scale_colour_gradientn(colours = c("green","red", "black"), aesthetics = c("colour", "fill"), na.value = "grey")+
  ylab("Highly pertubing miRNA nodes")+
  xlab("Patients")+
  facet_grid(rows = "situ")+
  theme( legend.title = element_blank(), panel.background = element_rect(fill = "gray93"), plot.margin = margin(0, 0, 0, 0, "cm"))

```