### 2021-02-19 multiple_samples_reproducible_codes.Rmd
example reprodcible codes for multiple samples

---
title: "Reproducible multiple sample script"
author:
- Selcen Ari
- Alper Yilmaz
date: "`r Sys.Date()`"
output: pdf_document
---

# Generating networks for multiple sample

This file contains an example of competing endogenous RNA analysis with ceRNAnetsim package.

```{r, include=FALSE, message=FALSE, warning=FALSE}
#required packages
library(tidyverse)
library(tidygraph)
library(readr)
library(biomaRt)
library(TCGAbiolinks)
library(ceRNAnetsim)
library(ggraph)
library(janitor)
library(SummarizedExperiment)
library(ceRNAnetsim)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# 1. Downloading data for analysis

```{r}

#query for Ovarian and Breast Carcinoma in TCGA. In this example TCGAbiolinks package was used.

query_gene_exp <- GDCquery(project = c("TCGA-BRCA", "TCGA-OV"),
                  data.category = "Transcriptome Profiling",
                  data.type = "Gene Expression Quantification", 
                  workflow.type = "HTSeq - Counts")

query_mirna_exp <- GDCquery(project = c("TCGA-BRCA", "TCGA-OV"),
                  data.category = "Transcriptome Profiling",
                  data.type = "Isoform Expression Quantification",
                  workflow.type = "BCGSC miRNA Profiling")

#finding samples which have both miRNA and gene expression dataset

getResults(query_gene_exp, cols = c("project","cases"))%>%
  mutate(barcode = substr(cases, 1,12))-> gene_exp_cases

getResults(query_mirna_exp, cols = c("project","cases"))%>%
  mutate(barcode = substr(cases, 1,12))-> mirna_exp_cases

# For demostration purposes we select 10 random patients per cancer project.
set.seed(1234)  
gene_exp_cases%>%
  inner_join(mirna_exp_cases, by = c("project", "barcode"))%>%
  dplyr::select(project, barcode)%>%
  distinct()%>%
  group_by(project)%>%
  sample_n(10)%>%
  pull(barcode)->selected_barcodes
```

## query for selected patient barcodes

```{r}
query_gene_exp_selected <- GDCquery(project = c("TCGA-BRCA", "TCGA-OV"),
                  data.category = "Transcriptome Profiling",
                  data.type = "Gene Expression Quantification", 
                  workflow.type = "HTSeq - Counts",
                  barcode = selected_barcodes)

query_mirna_exp_selected <- GDCquery(project = c("TCGA-BRCA", "TCGA-OV"),
                  data.category = "Transcriptome Profiling",
                  data.type = "Isoform Expression Quantification",
                  workflow.type = "BCGSC miRNA Profiling",
                  barcode = selected_barcodes)
```

## Downloading and preparing data

```{r}
GDCdownload(query_gene_exp_selected)
GDCdownload(query_mirna_exp_selected)

gene_exps <- GDCprepare(query_gene_exp_selected)
mirna_exps <- GDCprepare(query_mirna_exp_selected)

```

## Preparing gene expression dataset

```{r}
#preparation of gene expression dataset

as.data.frame(assay(gene_exps))%>%
  mutate(ensembl_gene_id = rownames(.))%>%
  dplyr::left_join(as.data.frame(rowData(gene_exps)), 
                    by = "ensembl_gene_id")%>%
  dplyr::select(Ensembl_Gene_Id = ensembl_gene_id, external_gene_name, 1:21)->gene_exp_to_be_analyzed

#fixing sample ids
cases <- which(substring(names(gene_exp_to_be_analyzed),1,4) %in% "TCGA")
names(gene_exp_to_be_analyzed)[cases] <- substr(names(gene_exp_to_be_analyzed)[cases], 1,12)

head(gene_exp_to_be_analyzed)
```

## Preparing miRNA expression dataset

```{r}

# miRNA expression dataset contains miRNAs with mirbase ids. So, firstly, mirbase_id_coversion dataset is downloaded and processed.

mirbase_url <- "ftp://mirbase.org/pub/mirbase/21/genomes/hsa.gff3"

read_tsv(mirbase_url, comment = "#", col_names = FALSE) %>%
  dplyr::select(mirna_type= X3, definition = X9)%>%
  filter(!endsWith(mirna_type, "primary_transcript"))%>%
  tidyr::separate(definition, c("ID", "Alias", "Name", "Derivated"), sep = ";")%>%
  dplyr::select(Alias, Name)%>%
  tidyr::separate(Alias, c("trash1", "ID"), sep = "=")%>%
  tidyr::separate(Name, c("trash2", "Name"), sep = "=")%>%
  dplyr::select(-trash1, -trash2)-> mirbase_id_conv

# preparation of miRNA expression dataset. 
# We used miRBase (Version 21) to obtain miRBase id (like MIMAT0000) for each mature isoform and 
# aggregated readr per million for each isoform.

mirna_exps%>%
  as.data.frame()%>%
  dplyr::select(miRNA_ID, 
                read_count, 
                reads_per_million_miRNA_mapped, 
                miRNA_region, 
                barcode)%>%
  dplyr::filter(startsWith(miRNA_region, "mature"))%>%
  dplyr::mutate(mirbase_id =str_remove(miRNA_region, "mature,"))%>%
  dplyr::select(-miRNA_region)%>%
  dplyr::inner_join(mirbase_id_conv, 
                    by = c("mirbase_id"="ID"))%>%
  dplyr::group_by(Name, barcode)%>%
  mutate(read_count= sum(read_count), 
         reads_per_million_miRNA_mapped = sum(reads_per_million_miRNA_mapped))%>%
  dplyr::ungroup()%>%
  dplyr::select(miRNA = Name, reads_per_million_miRNA_mapped, barcode)%>% # reads_per_million_miRNA_mapped used for ceRNA analysis (optionally)
  distinct()%>%
  tidyr::pivot_wider(names_from = "barcode", values_from = "reads_per_million_miRNA_mapped")-> mirna_exp_to_be_analyzed

#fixing sample ids
cases <- which(substring(names(mirna_exp_to_be_analyzed),1,4) %in% "TCGA")
names(mirna_exp_to_be_analyzed)[cases] <- substr(names(mirna_exp_to_be_analyzed)[cases], 1,12)

head(mirna_exp_to_be_analyzed)

```

# Performing competing endogenous RNA (ceRNAnetsim) analysis

Firstly, miRNA:gene pair dataset must be defined. Here, dataset of miRNA:gene pairs which were obtained from high-throughput experimental studies is used as an example. Note that in manuscript SPONGE analysis was preformed to refine miRNA:gene pairs. In this demonstration we omitted SPONGE analysis and use bare

```{r}

experimentalmirnagene <- readRDS("data/experimentalmirnagene.RDS")

graph_list = list()

for(i in selected_barcodes){

current_network <- experimentalmirnagene%>%
  right_join(dplyr::select(gene_exp_to_be_analyzed, Ensembl_Gene_Id, i), by="Ensembl_Gene_Id")%>%
  right_join(dplyr::select(mirna_exp_to_be_analyzed, miRNA, i), by ="miRNA", suffix= c("Gene_expression", "miRNA_expression"))%>%
  dplyr::select(Ensembl_Gene_Id, miRNA, Gene_expression= paste0(i, "Gene_expression"), miRNA_expression = paste0(i, "miRNA_expression"))%>%
  filter(!is.na(Gene_expression), !is.na(miRNA_expression))%>% 
  filter(Gene_expression != 0, miRNA_expression != 0)%>%
  priming_graph(competing_count = Gene_expression, 
                miRNA_count = miRNA_expression)

graph_list[[i]] <- current_network

}
```

Now we have miRNA:gene networks for 20 patients in a list. So any of them can be used with various functions provided by ceRNAnetsim package. Below we just printing out network of a single patient.

```{r}
graph_list$`TCGA-S3-AA17`
```